Lensing in the Schwarzschild Spacetime Applicable to Neutron Stars

### 2023-03-20 src/tex/ms.tex

>>>>

\section{Calculating the time delay}
%
\begin{linenomath}\begin{align}
    \label{eq:geodesic}
    ds^2 = g_{\mu\nu}dx^\mu dx^\nu = -(1-\frac{2}{r})dt^2 + \frac{1}{1-\frac{2}{r}}dr^2 + r^2(d\theta^2+\sin^2 \theta d\phi^2)
\end{align}\end{linenomath}
%
\begin{linenomath}\begin{align}
    \label{eq:momentum}
    p^\mu = \frac{d\chi^\mu(\lambda)}{d\lambda}
\end{align}\end{linenomath}
%

In order to calculate the time delay, we need to calculate $\frac{d\chi^t(\lambda)}{d\lambda}$ and $\frac{d\chi^r(\lambda)}{d\lambda}$. Then, the time delay
can be calculated as 
%
\begin{linenomath}\begin{align}
    \label{eq:timedelay}
    \Delta t = \int_{r_{NS}}^{\infty} dr \left(\frac{d\chi^t}{d\chi^r} - \frac{d\chi^t}{d\chi^r}\vert_{b=0} \right)
\end{align}\end{linenomath}
%

Let's calculate $\frac{d\chi^t(\lambda)}{d\lambda}$ and $\frac{d\chi^r(\lambda)}{d\lambda}$ assuming $\theta=\pi/2$ and therefore $p^\theta = 0$. The we have
in null geodesic:
%
\begin{linenomath}\begin{align}
    \label{eq:nullgeo}
    -(1-\frac{2}{r})(p^t)^2 + \frac{1}{1-\frac{2}{r}}(p^r)^2 + r^2\sin^2 \theta (p^\phi)^2 = 0
\end{align}\end{linenomath}
%
Solving eq. \ref{eq:nullgeo} leads to $p^t = \frac{d\chi^t(\lambda)}{d\lambda} = \frac{e}{1-\frac{2}{r}}$ and 
%
\begin{linenomath}\begin{align}
    \label{eq:pr-squared}
    \left(\frac{d\chi^r(\lambda)}{d\lambda}\right)^2 = \left(\frac{e^2}{1-\frac{2}{r}} - \frac{l}{r^2}\right) \left(1-\frac{2}{r}\right)
\end{align}\end{linenomath}
%
which, in terms of $b$ becomes
%
\begin{linenomath}\begin{align}
    \label{eq:pr-b}
    \frac{d\chi^r(\lambda)}{d\lambda} = e^2\sqrt{\left(\frac{1}{1-\frac{2}{r}} - \frac{b}{r^2}\right) \left(1-\frac{2}{r}\right)}
\end{align}\end{linenomath}
%
